---
title: "Red Wine Quality Exploration"
author: "Holly Moore"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    # theme of html document
    # theme of code highlight                                 
    # table of contents
    theme       : cosmo            # "default", "cerulean", "journal",
                                    # "flatly", "readable", "spacelab",
                                    # "united", "cosmo", "lumen", "paper", 
                                    # "sandstone", "simplex", "yeti"
    highlight   : tango          # "default", "tango", "pygments",
                                    # "kate",  "monochrome", "espresso",
                                    # "zenburn", "haddock", "textmate"
    toc         : true              # get table of content
    toc_depth   : 3
    toc_float   : true
    code_folding: hide
    number_sections: yes
    fig_caption: yes
    df_print    : paged

---

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# install.packages("knitr", dependencies = T)
# install.packages("dplyr", dependencies = T)
# install.packages("ggplot2", dependencies = T) 
# install.packages("gridExtra", dependencies = T)
# install.packages("ggcorrplot", dependencies = T)
# install.packages("PerformanceAnalytics", dependencies = T)

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(knitr)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(ggcorrplot)
library(PerformanceAnalytics)

# knitr: Suppress code/messages/warnings 
#        Set default plot options and center them
knitr::opts_chunk$set(fig.width=9,fig.height=5,fig.path='Figs/',
                      fig.align='center',tidy=TRUE,
                      echo=FALSE,warning=FALSE,message=FALSE)

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
# Load the Data
pf <- read.csv('wineQualityReds.csv')
```

The dataset used in this exploration contains 1,599 red wines. It has 11 
variables on the chemical properties of the wines. A minimum of three wine 
experts rated the quality of each wine, and they provided a rating between 
0 and 10 (0 being very bad, and 10 excellent). We will explore which chemical 
properties influence the quality of red wines.

# Univariated Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_Structure}
#glimpse(pf)
dim(pf)
str(pf)
```

Above, struture of the dataset

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_Summary}
summary(pf)
```

Above: summary of the dataset. The dataset contains 1599 rows and 13 columns. 
The first column labeled 'X' is simply row numbers, and I have chosen to 
exclude it from our dataset.  

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_Remove_colX}
pf <- subset(pf, select = -c(X))
names(pf)
```

That leaves us with 12 variables to explore and 1599 observations.

## Plot variables

```{r echo=FALSE, message=FALSE, warning=FALSE, Quick_Plot_Functions}
# We will be plotting a lot of variables. I decided to write a couple of 
# funtions to produce simple 1 variable plots. This will speed up this process 
# by eliminate repetitive code. I could have added nice color, but the 
# intructions say these should be quick and dirty.

f_ggplot1 <- function(fvariable, fmin, fmax, fstep, flabel, fbin)
  ggplot(aes(x = fvariable), data = pf) +
  #geom_histogram(binwidth = fbin, color = 'black', fill = '#C0C0C0') +
  geom_histogram(binwidth = fbin) +
  xlab(flabel) +
  scale_x_continuous(breaks = seq(fmin,fmax,fstep)) 

f_log_ggplot1 <- function(fvariable, fmin, fmax, fstep, flabel, fbin)
  ggplot(aes(x = fvariable), data = pf) +
  #geom_histogram(binwidth = fbin, color = 'black', fill = '#606060') +
  geom_histogram(binwidth = fbin) +
  labs(x=flabel) +
  xlab(flabel) +
  scale_x_log10(breaks = seq(fmin, fmax, fstep))
```

### Acidity Levels (Fixed, Volatile, & Citric)

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_Acidity}
pfa <- f_ggplot1(pf$fixed.acidity, 0, 16, 0.5, 'fixed.acidity', 0.5)
pva <- f_ggplot1(pf$volatile.acidity, 0, 1.75, 0.1, 'volatile.acidity', 0.05)
pca <- f_ggplot1(pf$citric.acid, 0, 1, 0.05, 'citric.acid', 0.01)

pfa
summary(pf$fixed.acidity)
pva
summary(pf$volatile.acidity)
pca
summary(pf$citric.acid)

grid.arrange(pfa, pva, pca, ncol = 1) #comparison of all 3 together
```

Looking at the acidity levels... 

Fixed acidity ranges from 4.6 to 15.9. It has a reasonably normal distribution, 
skewed slightly right, with a median 7.9 and mean of 8.32.

Volatile acidity ranges from .12 to 1.58. The distribution is also fairly 
normal, would be considered bimodal with peaks at 0.4 and 0.6, and 
slightly skewed right. The median and the mean are almost the same 
(median 0.52, mean 0.5278).

Citric acid ranges from 0 to 1, with most observations falling at 
zero, and another peak at 0.49. There is minimal citric acid in red wine.

### Residual Sugar Level

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_Sugar}
f_ggplot1(pf$residual.sugar, 0.9, 15.5, 1, 'residual.sugar', 0.5)
f_log_ggplot1(pf$residual.sugar, 0, 15, 1, 'residual.sugar (scale log10)', .05)

summary(pf$residual.sugar)
```

Residual sugar levels range from 0.9 to 15.5, with a median 2.2 and a mean of 
2.539. The distribution is heavily skewed right, with most observations 
falling under 3. Since we have a very long tail, the bottom graph is 
transformed using a log scale 10 on the x-axis to get a better look.

### Chloride level

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_Chloride}
f_ggplot1(pf$chlorides, 0.012, 0.611, 0.1, 'chlorides', .02)
f_log_ggplot1(pf$chlorides, 0.012, 0.611, 0.05, 'chlorides (scale log10)', .05)

summary(pf$chlorides)
```

The Chloride levels range from 0.012 to 0.611 and are heavily skewed right, 
with 75% of the observations at or below 0.09. Since we have a long tail, the 
bottom graph (scale x log 10), zooms in so we can see what is going on. On 
the transformed plot (excluding observations above 0.20), the distribution 
appears normal.

### Free Sulfur Dioxide Levels

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_Free_Sulfur}
fsd <- 'free.sulfur.dioxide'

f_ggplot1(pf$free.sulfur.dioxide, 0, 73, 5, fsd, 2)
f_log_ggplot1(pf$free.sulfur.dioxide, 0, 73, 5, fsd, 0.03)

summary(pf$free.sulfur.dioxide)
table(pf$free.sulfur.dioxide)
```

Free sulfur dioxide has a wide range from 1 to 72, but 75% of observations 
fall at 21 and below. The median is 14, and the median is 15.87. The 
distribution is heavily skewed right like many of the other variables we 
have looked at. Due to the long tail, logscale was used on the x-axis, 
allowing us to see a sharp peak 6. After seeing the peak at 6, I ran the table 
command so we could see how many of each number we were dealing with. There 
are 138 observations with number 6. 

### Total Sulfur Dioxide Levels

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_Total_Sulfur}
tsd <- 'total.sulfur.dioxide'

f_ggplot1(pf$total.sulfur.dioxide, 0, 300, 20, tsd, 5)
f_log_ggplot1(pf$total.sulfur.dioxide, 0, 300, 20, tsd, .05)

summary(pf$total.sulfur.dioxide)
```

Total sulfur dioxide has an even wider range, from 6 to 289, but 75% of 
observations fall at 62 and below. The median is 38, and the median is 46.47. 
The distribution is heavily skewed right. We used logscale to transform the 
x-axis, giving us a better look at the data. The transformed data has a 
normal distribution peaking between 40-50.

### Density

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_Total_Density}
f_ggplot1(pf$density, min(pf$density), max(pf$density), .002, 'density', .0006)


summary(pf$density)
```

The density of wine has a minimal variance range, from 0.9901 to 1.0037 
(difference of only 0.0136). The minimal variance is not surprising since 
all wine has the consistency of water, which is 1g/cm3. The distribution is 
normal with the median and the mean being almost the same (median 0.9967 & 
mean 0.9968).

### pH Levels

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_Total_pH}
f_ggplot1(pf$pH, min(pf$pH), max(pf$pH), 0.1, 'pH', 0.1)

summary(pf$pH)
```

The pH levels also have a small variance range, from 2.74 to 4.01 
(difference of 1.27). The distribution is normal, almost symmetrical, with 
the median and the mean being almost the same (median 3.310 & mean 3.311).

### Sulphate Levels

```{r echo=FALSE,message=FALSE, warning=FALSE, Univariate_Plots_Total_Sulphates}
f_ggplot1(pf$sulphates, 0.3, 2, 0.1, 'sulphates', 0.1)

summary(pf$sulphates)
```

The sulphate levels range from 0.33 to 2.0, but 75% of all observations fall 
at 0.73 and below. The median is 0.62, and mean 0.6581. The distribution is 
right-skewed with some outliners on the far left.

### Alcohol Level

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_Alcohol}
f_ggplot1(pf$alcohol, 8, 15, 0.5, 'alcohol', 1/4)

summary(pf$alcohol)
```

The alcohol level ranges from 8.4 to 14.9. The mean and the median are almost 
the same (median 10.2, mean 10.42). The distribution is right-skewed with the 
peak at 9.5. The highest rating of 14.9 looks like an outliner.

### Quality Ratings

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_Quality}
ggplot(aes(x = quality), data = pf) +
  geom_bar() +
  scale_x_continuous(breaks = 3:8)

summary(pf$quality)
table(pf$quality)
```

The quality rating scale goes from 0 to 10, with 10 being the best. However, in 
this dataset, our ratings go from 3 to 8 (8 being the best). The distribution is 
normal, peaking between five & six. Most of the observations are between 5 & 7.

# Univariate Analysis

### What is the structure of your dataset?
This dataset contains 1,599 observations and has twelve variables (11 of which 
are chemical properties of the wines). The variables are: fixed.acidity, 
volatile.acidity, citric.acid, residual.sugar, chlorides, free.sulfur.dioxide, 
total.sulfur.dioxide, density, pH, sulphates, alcohol, and quality. 

### What is/are the main feature(s) of interest in your dataset? 
The main item of interest is wine quality. We are exploring the chemical 
properties to determine which ones influence the quality of red wines.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

Since we are interested in red wine quality, we want to locate the variables 
that affect quality. Right now, I am interested in the variables that correlate 
with quality. 

### Did you create any new variables from existing variables in the dataset?
No.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form 
of the data? If so, why did you do this?

The dataset came nice and tidy. While making the initial variable plots, 
adjusting the bin size was about all that needed to be done. And four of the 
variables (residual.sugar, chlorides, free.sulfur.dioxide, 
total.sulfur.dioxide) had long tails. In order to zoom in and get a 
better picture of the distribution, I used scale_x_log10 on the x-axis to 
transform the data. 

# Bivariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots}
# http://jamesmarquezportfolio.com/correlation_matrices_in_r.html
# http://www.sthda.com/english/wiki/ggcorrplot-visualization-of-a-correlation-
# matrix-using-ggplot2
# https://cran.r-project.org/web/packages/ggcorrplot/ggcorrplot.pdf
# library("PerformanceAnalytics")
# library(ggcorrplot)
chart.Correlation(pf, histogram=TRUE, pch=19)
ggcorrplot(cor(pf), hc.order=TRUE, type='lower', lab = TRUE, insig = "blank", 
           lab_size=3)
```

Above, we have two different correlation matrix plots. They make it easy to see 
the strength of the relationships between the variables. 

In the first matrix plot, the higher the correlation, the bigger the font used 
for the correlation strength. \
For fun, I threw in the second matrix plot. In this matrix, the strength of the 
relationship is represented by color, the darker the shade 
(either red or purple), the stronger the relationship. 

**Correlations with Quality** \
Although some of these correlations are relatively weak, they are still listed 
here.
Variables with the strongest correlation to quality are: \
 0.48 alcohol \
-0.39 volatile acidity \
 0.25 sulphates \
 0.23 citric.acid \

As for the relationships between the other supporting variables, there are 
quite a few interesting correlations we can see (both positive and negative).

**Negative correlations** \
-0.68 pH vs. fixed.acidity \
-0.55 volatile.acidity vs. citric.acid \
-0.54 pH vs. citric.acid \
-0.50 alcohol vs. density \

**Positive correlations** \
 0.67 fixed.acidity vs. citric.acid \
 0.67 fixed.acidity vs. density \
 0.67 total.sulfur.dioxide vs. free.sulfure.dioxide \
 0.37 sulphates vs. chlorides \
 0.36 residual.sugar vs. density \
 
I am surprised to see only a few variables correlate to wine quality, I 
expected more. However, there are still plenty of correlations to explore. 

Now we will plot some of the correlations!

## Negative Correlations

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivarate_Plots_functions}
# Function that returns geom_point plot with 2 variables
f_gg_point <- function(fvarY, fvarX, fbin, falpha, fmethod, labelY, labelX)
  ggplot(pf, aes(y = fvarY, x = fvarX), binwidth = fbin) +
  geom_point(alpha = falpha) + #example 1/20 sets alpha level 20 per point
  geom_smooth(se = FALSE, method = fmethod) +
  xlab(labelX) +
  ylab(labelY) 

# Noise in conditional means (from lesson 4), smooth the conditional means
f_gg_line <- function(fvarY, fvarX)
  ggplot(aes(x = fvarX, y = fvarY), data=pf) +
  geom_line(stat = 'summary', fun.y = mean)


f_gg_smooth <- function(fvarY, fvarX, flimY, flimX, falpha)
  ggplot(pf, aes(x = fvarX, y = fvarY)) +
  stat_smooth(methos = 'lm') +
  scale_x_continuous(lim = c(flimX, quantile(fvarX, .99))) +
  scale_y_continuous(lim = c(flimY, quantile(fvarY, .99))) +
  geom_point(alpha = falpha, fill = I('#F79420'), color = I('black'), 
             shape = 21)

# Function for Quality - create a point plot
f_quality_point <- function(fvarY, labelY)
  ggplot(pf, aes(y = fvarY, x = quality)) +
  geom_point(aes(colour=factor(quality))) +
  xlab('Quality') +
  ylab(labelY) 

# Function for Quality - create a boxplot
f_quality_boxplot <- function(fvarY, labelY)
  ggplot(pf, aes(y = fvarY, x = factor(quality))) +
  geom_boxplot() +
  stat_summary(fun.y=mean, geom="point", shape=4) +
  xlab('Quality') +
  ylab(labelY) 

# Function for Quality - create a jitter plot
f_quality_jitter <- function(fvarY, labelY, falpha)
  ggplot(pf, aes(x = quality, y = fvarY, color = factor(quality))) +
  geom_jitter(alpha = falpha) +
  xlab('Quality') +
  ylab(labelY) 
```


### pH vs. Fixed Acidity

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_PH_FA}
f_gg_point(pf$pH, pf$fixed.acidity, 1, .1, 'lm', 'pH', 'fixed.acidity')
f_gg_line(pf$pH, pf$fixed.acidity)

cor.test(pf$pH, pf$fixed.acidity)
summary(pf$pH)
summary(pf$fixed.acidity)
```

The variables pH and fixed acidity have the highest negative correlation of 
all the variables (corr -0.68). The plot shows that as pH increases, the level 
of fixed acidity decreases. We made a point and line plot, both are interesting
to look at.

### Volatile Acidity vs. Citric Acid

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_VA_CA}
f_gg_point(pf$volatile.acidity, pf$citric.acid, 1, .15, 'lm', 
           'volatile.acidity', 'citric.acid')
f_gg_smooth(pf$volatile.acidity, pf$citric.acid, .3, 0, .25)

cor.test(pf$volatile.acidity, pf$citric.acid)
summary(pf$volatile.acidity)
summary(pf$citric.acid)
```

These two variables, volatile acidity and citric acid, have the 2nd highest 
negative correlation (corr -0.55). So, it is not surprising that the plot looks 
similar to the previous one. As volatile acidity increases, citric acid 
decreases. On the second plot, we removed the outliners to get a better 
view of the distribution.

### pH vs. Citric Acid

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_pH_CA}
f_gg_point(pf$pH, pf$citric.acid, 1, .50, 'lm', 'pH', 'citric.acid')
f_gg_point(pf$pH, pf$citric.acid, 1, .05, 'lm', 'pH', 'citric.acid')
f_gg_smooth(pf$pH, pf$citric.acid, 3, 0, .15)

summary(pf$citric.acid)
summary(pf$pH)
cor.test(pf$pH, pf$citric.acid)
```

Above is the last negative correlation we will do for variables outside of 
quality. Again, as citric acid increases, pH decreases (corr -0.54). We do 
have an outliner at 1. I made two different plots, one with alpha at .5 and 
the other with alpha at .05. The plot with .05 makes it easy to see where 
we have spikes of citric acid. On the final plot, we limited the range, 
allowing us a better view of the distribution.

## Positive correlations

### Fixed Acidity and Density

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_FixedAcidity&Density}
f_gg_point(pf$density, pf$fixed.acidity, 1, .25, 'lm', 
           'density', 'fixed.acidity')
f_gg_line(pf$density, pf$fixed.acidity)

summary(pf$density)
summary(pf$fixed.acidity)
cor.test(pf$density, pf$fixed.acidity)
```

Above is the first positive correlation we have explored (corr is .67). Making
it clear to see that as fixed acidity increases, so does density. The line 
chart does a nice job of showing the positive pattern, while pointing out the
outliners.

### Free Sulfer Dioxide vs Total Sulfer Dioxide

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_FSD_TSD}
f_gg_point(pf$free.sulfur.dioxide, pf$total.sulfur.dioxide, 
           1, .25, 'lm', 'free.sulfur.dioxide', 'total.sulfur.dioxide')

ggplot(pf, aes(y = free.sulfur.dioxide, x = total.sulfur.dioxide)) +
  geom_point(alpha = .05, position = position_jitter(h = 0)) +
  geom_smooth(se = FALSE, method = "lm") +
  xlim(0, 200) +   
  ylim(0, 60)

cor.test(pf$free.sulfur.dioxide, pf$total.sulfur.dioxide)
summary(pf$free.sulfur.dioxide)
summary(pf$total.sulfur.dioxide)
```

The above plot is funny; it looks like a megaphone. These two variables have a 
strong correlation (.67), in a three-way tie for the strongest 
positive correlation in this dataset. Since they are both a form of sulfur 
dioxide, it is not surprising they have a strong correlation. The second
plot limits both the x an y axis, giving us a better view of the distribution.

## Correlations with Quality!

### Quality vs Alcohol

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Quality_Alcohol}
#f_quality_point(pf$alcohol, 'alcohol')
f_quality_jitter(pf$alcohol, 'alcohol', .25)
f_quality_boxplot(pf$alcohol, 'alcohol')

cor.test(pf$alcohol, pf$quality)
print('alcohol summary')
summary(pf$alcohol) 
by(pf$alcohol, pf$quality, summary)
```

We are finally looking at the correlations with quality. This is interesting, 
but not surprising. The level of alcohol correlates with red wine 
quality at 0.48. Looking at these plots, you can see the gradual 
increase in quality as alcohol level increases. It is fun to look at 
these different plots; the boxplot especially does a great job visualizing the 
positive correlation. I do notice a slight decrease in alcohol level in the 
wine rated quality level 5. I do wonder what caused this decrease?

### Quality vs Volatile Acidity

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Quality&VA}
#f_quality_point(pf$volatile.acidity, 'volatile acidity')
f_quality_jitter(pf$volatile.acidity, 'volatile acidity', .25)
f_quality_boxplot(pf$volatile.acidity, 'volatile acidity')

cor.test(pf$volatile.acidity, pf$quality)
print('volatile.acidity summary')
summary(pf$volatile.acidity) 
by(pf$volatile.acidity, pf$quality, summary)
```

Now we have a negative correlation between red wine quality and volatile 
acidity (corr -0.39). This is the next highest correlation for quality. 
On all three plots, you can easily see the negative relationship. As volatile 
acidity decreases, the quality also decreases. Again, the boxplot does an 
excellent job of giving a visual on the correlation direction. 

### Quality vs Sulphates

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Quality&Sulphates}
#f_quality_point(pf$sulphates, 'sulphates')
f_quality_jitter(pf$sulphates, 'sulphates', .25)
f_quality_boxplot(pf$sulphates, 'sulphates')

cor.test(pf$sulphates, pf$quality)
print('sulphates summary')
summary(pf$sulphates) 
by(pf$sulphates, pf$quality, summary)
```

Above is a relatively weak correlation (0.25 for sulphates vs quality), but 
since our primary interest is quality, I wanted to explore as many variables  
as possible (that correlate with quality). As expected, with a weak 
correlation, there is only a slight difference between the levels. But it is 
a positive correlation, as the sulphates increase, so does the wine's quality.
I thought it was interesting that the highest quality level (8), has a 
minimum sulphate level of 0.63, close to the median level for the other 
quality ratings.

### Quality vs Citric Acid

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Quality&CA}
#f_quality_point(pf$citric.acid, 'citric acid')
f_quality_jitter(pf$citric.acid, 'citric acid', .25)
f_quality_boxplot(pf$citric.acid, 'citric acid')

cor.test(pf$citric.acid, pf$quality) 
print('citric.acid summary')
summary(pf$citric.acid) 
by(pf$citric.acid, pf$quality, summary)
```

Above is the last and also the lowest quality correlation that we are going to 
explore. Looking at the first plot, it appears that quality levels 5 and 6 
contain just about every level of citric acid. However, taking a look at the 
jitter plot, you can get a better picture. However, the exciting plot here 
is the boxplot. On the boxplot, you can see an apparent positive correlation 
between citric acid and quality. As the levels of citric acid increase, so does 
the quality. I did not expect to see this!

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in 
the dataset?

We have just finished plotting various positive and negative correlations 
between variables in the dataset. We used several different plot styles to give 
a different perspective on the data. We are most interested in red wine quality,
so we examined 4 different variables against quality. The strongest correlation 
we found towards quality was with alcohol (the more alcohol, the higher quality 
of the wine).

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

The three highest positive correlations all had corr 0.67. \
These correlations were: \
1) fixed acidity vs citric acid \
2) fixed acidity vs density \
3) total sulfur dioxide vs free sulfur dioxide \
The top and the bottom one (fixed acidity vs citric acid and total sulfur 
dioxide vs free sulfur dioxide) were not surprising since they are 
both acids and the other both sulfur dioxide.

### What was the strongest relationship you found?

The relationship between pH and fixed acidity has the strongest correlation 
in the dataset (corr -0.68).

# Multivariate Plots Section

### Fixed acidity vs Density with Quality

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots1}
ggplot(pf, aes(x = fixed.acidity, y = density)) +
  #geom_point(alpha = .25) +
  geom_point(aes(colour=factor(quality)), alpha = .4) +
# https://stackoverflow.com/questions/29554796/meaning-of-band-width-in-
# ggplot-geom-smooth-lm
  geom_smooth(se = FALSE, method = "lm")
```

I was interested in adding quality to one of the highest non-quality 
correlations we did. Fixed acidity and density have a positive correlation
of 0.67. This result was very surprising! As density increases, so does 
fixed acidity. However, the highest quality rated wines, stay at the bottom
of the results (the lowest density, the lowest fixed acidity). 

### Citric acid vs pH with Quality

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots2}
ggplot(aes(citric.acid, pH, color = factor(quality)), data = pf) +
  geom_jitter(size = 2, alpha = 1/2) +
  scale_color_brewer(type='qual')
#  theme_classic()
```

We used the correlation between pH and citric acid (-0.54) and added quality 
to it for our second multivariate plot. The result was not helpful. Our 
top-quality rating of '8' appears at all levels on this chart. It looks like 
there is no correlation between quality and these two variables.

### Alcohol vs Volatile Acidity with Quality

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots3}
ggplot(aes(x=volatile.acidity, y=alcohol, color=factor(quality)), data=pf) + 
    geom_point(alpha = 1) + 
    scale_color_brewer(type='div', palette=2)

ggplot(aes( alcohol, volatile.acidity,color = factor(quality)), data = pf) +
  geom_jitter(size = 2, alpha = .5) +
  scale_color_brewer(type='qual')
  #theme_light()
```

Here we have the two variables with the highest correlation with quality 
and then added quality to them. These results were not surprising as we already
knew higher quality red wines had higher levels of alcohol and lower levels 
of volatile acidity. However, they are still interesting to look at in the 
plots. I could not decide which way I wanted to plot these, with alcohol on 
the x-axis or on the y-axis, so I did one of each. The first plot turned out to
be my favorite, so this will be in my final plot section. 

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots4}
ggplot(aes(citric.acid, sulphates, color = factor(quality)), data = pf) +
  geom_jitter(size = 2, alpha = .5) +
  theme_light()
```

The remaining two variables we looked at, with correlations to quality, are 
Sulphates and citric acid. So we used those two variables in this plot and 
added quality to them. This turned out a little interesting. Because most 
quality levels are between 5-6, this plot is not very clear. Nevertheless, 
it is noticeable that quality levels 7 & 8 seem to fall between sulphate levels
.4 to .8 (roughly). The sulphate outliners drop back down to average quality 
wine. Citric acid levels do not seem to matter as much as we can see high levels 
throughout the spread of citric acid. 

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of 
looking at your feature(s) of interest?

In our plots, we found that high-quality wines contain higher levels of 
alcohol, sulphates, and citric acid. And lower levels of volatile acidity. 

### Were there any interesting or surprising interactions between features?

Yes, the plot for fixed acidity vs density, with quality as color. It was 
very interesting to see that as the levels of each climbed, the higher wine
quality stayed to the bottom (with lower density, and lower fixed acidity).

### OPTIONAL: Did you create any models with your dataset? Discuss the \

No models were made.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_One}
f_quality_jitter(pf$alcohol, 'Alcohol Level', .3) +
  ggtitle('       Red Wine Quality by Alcohol Level') +
  xlab('Quality Level') +
  theme_light(15)
```

### Description One

I like the colors on this plot, and I like how you can easily see the quality 
level rise as the alcohol level increases. The alcohol level was the variable 
with the highest correlation with quality (corr 0.48). I did find it surprising 
and disappointing that most of the quality ratings are between 5-7. 95% of the 
ratings are between 5-6. So, perhaps if our sample size included more wines9, 
we would have more ratings fall on the scale's low and high end.

### Plot Two
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Two}
  ggplot(data = pf, aes(y = citric.acid, x = factor(quality), 
                        color=factor(quality))) +
  geom_boxplot() +
  ggtitle('        Citric Acid levels in Red Wine Quality') + 
  xlab('Quality Level') +
  ylab('Citric Acid Level') +
  theme_light(14)
```

### Description Two

I selected this plot because I found it surprising. The correlation between 
quality and citric acid was  relatively weak, being 0.23, so I did not expect 
this boxplot to show a nice steady increase in quality, as the level of citric 
acid went up. I added some color onto it because I enjoy seeing the color. 
Even though we have a limited sample size, and the amounts of citric acid 
are minimal, it appears the level of citric acid used could play a role in the 
quality produced.

### Plot Three
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Three}
ggplot(aes(x=volatile.acidity, y=alcohol, color=factor(quality)), data=pf) + 
    geom_point(alpha = .70) + 
    scale_color_brewer(type='div', palette=2) +
    ggtitle('     Red Wine Quality Level: Alcohol vs Volatile Acidity Levels')+   
    xlab('Level of Volatile Acidity') +  
    ylab('Percentage of Alcohol') +
    theme_light()
```

### Description Three

I saved this plot for last because it is my favorite. This plot has wine 
quality along with the two highest correlations to wine quality. Alcohol has 
a correlation of 0.48, and volatile acidity's correlation is -0.39. I  
like to see how these two variables come together, matched up with quality. 
With the greens and pinks, it makes it stand out, higher levels of alcohol 
along with low levels of volatile acidity equal better red wine quality.  

------

# Reflection

In this dataset, we examined 1,599 different red wines. The goal was to 
determine which variables contributed to the quality of the wine. I started 
by spending time looking at each variable individually. 

Then I checked the correlation between the variables. It was surprising to 
find that the highest correlation was only -0.68. I had expected alcohol level 
to have a higher correlation, but then I suppose that would make it too easy to 
produce quality wine. If the alcohol level were the key, wine makers could9 
increase the alcohol level and be done with it, but nothing is ever that 
simple. After that, then I spent some time plotting the correlations of 
interest. I had hoped there would be more variables correlate with quality, but 
I used what we had. The most surprising plot for me was the correlation between 
quality and citric acid. It had a very weak correlation, and citric acid levels 
are very low in wine in general. I did not expect to see much of anything when 
I plotted these two variables. However, instead, we saw a very definite 
increase in quality as the citric acid level increased.

In the multivariate plots, we were able to get a look at three variables 
together. My favorite plot was putting alcohol, volatile acidity, and quality 
together. I had examined these variables separately, but bringing them together 
painted a nice picture of red wine quality (higher alcohol levels, along with 
lower volatile acidity = better red wine quality). 

I selected the smallest dataset to use for this project. Now I wonder how this 
would have gone had I selected the dataset with a larger sample size. I would 
have like to see the quality ratings on a larger sample size. I am curious if 
we would still see 95% of the wines falling between a quality rating of 5-7 
or getting more variety. Had there been more wines with the highest 
quality rating, perhaps it would have been easier to determine which chemical 
properties influence the quality of red wines.

This was my first time using Rstudio, and the options seem endless! I loved 
exploring different things and seeing all the different types of plots and 
analyses available in R. 